Examining background gene list issue in clusterProfiler - RNA-seq example

### 2024-09-16 analysis/analyse_rnaseq1.Rmd
savedata at end
added deepdive

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("fgsea")
  library("eulerr")
  library("gplots")
  library("parallel")
  library("RhpcBLASctl")
})

@@

@@

>>>>

suppressPackageStartupMessages({
  library("getDEE2")
  library("DESeq2")
  library("kableExtra")
  library("clusterProfiler")
  library("fgsea")
  library("eulerr")
  library("gplots")
  library("parallel")
  library("RhpcBLASctl")
  library("Hmisc")
})

@@

## Focus on dataset 1 down with HPO genesets

This extreme example will help us understand the bias better.

```{r,hpodn}

orig <- orabad(d[[1]],gs[[7]])
corr <- oragood(d[[1]],gs[[7]])
orig_dn <- orig[[2]]
corr_dn <- corr[[2]]
mg <- merge(orig_dn,corr_dn,by=0)

mg$size <- as.numeric(sapply(strsplit(mg$BgRatio.x,"/"),"[[",1))
quantiles <- quantile(mg$size, prob = seq(0, 1, length = 6), type = 5)
mg$levels <- cut2(mg$size, cuts = as.numeric(quantiles))
cols <- gsub("5","red",gsub("4","orange",gsub("3","yellow",gsub("2","green",gsub("1","blue",as.character(as.numeric(mg$levels)))))))

plot(-log10(mg$pvalue.x),-log10(mg$pvalue.y),
  ylim=c(0,7), xlim=c(0,7), pch=19,
  main="SRP128998 HPO down",
  xlab="-log10 p-value (original)", ylab="-log10 p-value (corrected)")

grid()
abline(coef = c(0,1))

points(-log10(mg$pvalue.x),-log10(mg$pvalue.y),
  col=cols, pch=19,cex=0.7)

pdf("fig1_HPOdn.pdf",width=7, height=7)
plot(-log10(mg$pvalue.x),-log10(mg$pvalue.y),
  ylim=c(0,7), xlim=c(0,7), pch=19,
  main="SRP128998 HPO down",
  xlab="-log10 p-value (original)", ylab="-log10 p-value (corrected)")

grid()
abline(coef = c(0,1))

points(-log10(mg$pvalue.x),-log10(mg$pvalue.y),
  col=cols, pch=19,cex=0.7)
dev.off()

@@

save.image("analyse_rnaseq1.Rdata")
